Supplementary Materials- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period

### 2021-01-26 analysis/paper/paper.Rmd
edit the paper rmd from intro to method

<!--burial study- network analysis-->
Burial analysis by archaeologists is an approach to understand past social structures through the study of the physical traces of mortuary practices. The material cultures and biological records of burial behaviors can represent the social ranking or identities of the deceased [@Binford1971; @Drennan2010; @Saxe1970]. Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003]. For example, mortuary practices, including burial forms, grave goods, and ritual behaviors that structure the material configuration of burials, can represent social relations between members of a community, allowing the inference of social structures [@Byrd1995; @Seikel2011]. Based on network science and graph theory, the recent development of social network analysis provides many new tools to visualize and analyze relationships in archaeological data [@Borgatti2009; @Brughmans2013]. Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013; @Collar2015]. A network can be viewed as a patterned aggregation that includes individual elements (i.e. individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data. The details about archaeological networks provided in Supplementary Online Materials. This paper will use a novel Bayesian approach on exponential random graph model for networks to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.

@@

<!-- archaeological background -->
Archaeological sites in northeastern Taiwan show evidence of imported prestige goods, such as ornaments and porcelains, in burials and some accumulation patterns in residential areas around the same time as European contact in the 17th century [@Chen2007; @LWandBM2020ornament]. The use of imported goods is also mentioned in historical records from the Spanish, indicative of the pursuit of prestige or wealth in those Indigenous societies [@Borao2009; @LiandWu2006]. For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006]. In addition, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006]. Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear. The use of trade items as prestige goods in local Indigenous communities might occur earlier before a European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016]. A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks [@Brumfiel1994; @Clark1994]. The observed uneven distribution may be explained as a result of social change from less social inequality to more social inequality when Indigenous societies in Northeastern Taiwan were involved in the complex trade network stimulated by the Europeans.

@@

<!-- archaeological specific case study -->
In this study, we explore the changes in social inequality in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007]. We assume that social changes at Kiwulan would be supported by chronological differences in the structure of burial networks. We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival. We ask: (1) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by analysis of burial networks? (3) what are the major variables affecting or forming the higher degree of popularity (e.g. a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period? By answering these questions, this study helps to expand the use of statistical modeling for burials in understanding the indirect effects of a colonial presence on Indigenous groups. This is important to understand reactions to indirect impacts of European colonization.

@@

<!-- ERGM in archaeology, and limitations --> 
Exponential random graph models first appeared in archaeology with @Brughmans2014, who studied  Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites. Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes. Similarly, @Amati2019 reconstructed three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore interaction mechanisms. These network models evaluated different hypotheses about the importance of three variables, proximity, inter-cultural items, and pottery types, on the presence of hub sites they observed. They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively. However, those studies also point out some limitations of ERGM, such as sensitivity to missing data and less ability to handle uncertainty. In addition, it is difficult in ERGM to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013]. A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian]. This is also termed "doubly intractable" since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015]. Model degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014]. One solution to these limitations is available by implementing ERGMs in a Bayesian framework.

@@

<!-- Bayesian inference for Exponential Random Graph Models --> 
Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporate prior information about the network configurations into the model and offer uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019]. In Bayesian analysis, posterior distribution is an updated probability distribution after combining the prior derived from previous data or assumptions and our data. The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011]. MCMC avoids doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results. This enable us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006]. By fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters. Posterior probability estimates the effect of ERGM parameters by looking at the posterior mean and 95% credible intervals, which can replace p-values for assessing a null hypothesis [@Caimo2017bayesian]. The typical criteria for interpreting the posteriors is that an odds ratio greater than one means a positive effect of a parameter, while odds ratios less than one represents negative effect. In addition, Bayesian approaches are also useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies. @Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on the average, 80% of the ties when a third of data missing.

@@

<!-- archaeological setting -->
We analyzed burial data collected from the published excavation reports, and the original fieldwork notes for the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD). The settlement experienced European colonial impacts in the early 17th century and a large wave of Chinese immigrants in the 19th century [@Chen2007]. Excavations revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007]. In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with _in-situ_ posts were also excavated. The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)). Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, indicating a well organized spatial arrangement. Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences. For example, @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society. However, @Hsieh2012 suggested a relatively egalitarian structure based on comparative analysis of the frequencies of all burial goods. She found that the burials with high value burial goods were usually associated with elders, which might indicate achieved, rather than inherited, status. One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organization with post-European social organization. Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period to the post-European period. The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well studied.

@@

```{r sample-size}
# number of each phase
library(tidyverse)
pre_E <- burial_three_period_age_number %>% filter(Phase == "pre") %>% pull()
post_E <- burial_three_period_age_number %>% filter(Phase == "post") %>% pull()
chi <- burial_three_period_age_number %>% filter(Phase == "chi") %>% pull()
diff_pre_post <- post_E - pre_E
```

<!-- chronology -->
To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`). Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [@LWandBM2020ornament; @Chen2007]. We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size. There are eight burials heavily disturbed by modern construction that are also excluded because we cannot determine their chronology. We reconstructed social relations between individuals by linking them according to similar prestige goods [cf. @Coward2013, pp. 252]. Therefore, we built networks where burials represent actors (nodes in the network) that are linked when they have the same prestige goods in common. The prestige goods we identified include agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments. Previous studies indicate that these items are considered as high-value across different archaeological contexts [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011]. Historical accounts also support these items as prestige goods [@Borao2009; @LiandWu2006].

@@

<!-- our hypothesis --> 
Our pericolonial impact hypothesis is that rare, high quality grave goods accumulated in only a few individuals' burials after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access to these goods. We can identify these high value trade goods from historical documents written by Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power. If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status). To test this prediction, we use ERGM in a Bayesian framework to examine the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan. By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.

@@

<!-- burial data and the definition of tie formation  --> 
We used burial goods in common to construct ties for representing a flow of good between burials. Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials). For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>28, about 10 percent), above average (12-28), and below average (<12), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship. The networks constructed based on this principles show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:figure-network-diagrams)).

>>>>

<!--burial study- network analysis-->
Burial analysis by archaeologists is an approach to understand past social structures through the study of the physical traces of mortuary practices. The material cultures and biological records of burial behaviors can represent the social ranking or identities of the deceased [@Binford1971; @Drennan2010; @Saxe1970]. Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003]. By studying burial forms, grave goods, and ritual behaviors that structure the material configuration of burials, we can infer social status and relations between members of a community that give an insight to social structures [@Byrd1995; @Seikel2011]. Based on network science and graph theory, the recent development of social network analysis provides many new tools to visualize and analyze relationships in archaeological data [@Borgatti2009; @Brughmans2013]. Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation [@Brandes2013; @Collar2015]. A network can be viewed as a patterned aggregation that includes individual elements (i.e. individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data. Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg. @Coward2013, pp. 252]. This paper will use a novel Bayesian approach on exponential random graph model for networks to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign presence on social organization.

@@

<!-- archaeological background -->
Archaeological sites in northeastern Taiwan show evidence of accumulation patterns of imported goods, such as ornaments and porcelains, in residential areas and the use as burial goods around the same time as European contacts in the 17th century [@Chen2007; @LWandBM2020ornament]. Imported goods were treated as prestige or wealth in those Indigenous societies that were commonly mentioned in Spanish historical records [@Borao2009; @LiandWu2006]. For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006]. In addition, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006]. Despite we observed the uneven distribution of imported goods in burials, the degree of differentiation across burials over time remain unclear. The use of imported items as prestige goods in local Indigenous communities might occur earlier before a European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016]. A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994]. The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade networks when Indigenous societies were involved in the complex trade network stimulated by the Europeans.

@@

<!-- archaeological specific case study -->
In this study, we explore social changes in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007]. We assume that social changes at Kiwulan would be supported by chronological differences in the structure of burial networks. We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival. We ask: (1) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by analysis of burial networks? (2) what are the major variables affecting or forming the higher degree of popularity (e.g. a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period? By answering these questions, this study helps to expand the use of statistical modeling for burials in understanding the indirect effects of a colonial presence on Indigenous groups. This is important to understand reactions to indirect impacts of European colonization.

@@

<!-- ERGM in archaeology, and limitations --> 
Exponential random graph models first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites. Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes. Similarly, @Amati2019 modeled three networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore interaction mechanisms, including proximity, inter-cultural items, and pottery types. By comparing hypothesized networks with the observed sites, they found that the presence of hub sites can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively. However, those studies also point out some limitations of ERGM, such as sensitivity to missing data and less ability to handle uncertainty. In addition, it is difficult in ERGM to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013]. A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian]. This is also termed "doubly intractable" since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015]. Model degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014]. One solution to these limitations is available by implementing ERGMs in a Bayesian framework.

@@

<!-- Bayesian inference for Exponential Random Graph Models --> 
Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective tools for network modeling by incorporating prior information about the network configurations to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020]. The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011]. MCMC avoids doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results. This enable us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006]. After fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporates our sample data and prior information derived from previous data or our assumptions through summary statistics from estimates of the ERGM parameters [@Caimo2015; @Nemmers2019]. This also allows uncertainty quantification. Posterior probability estimates the effect of ERGM parameters by looking at the posterior mean and 95% credible intervals, which can replace p-values for assessing a null hypothesis [@Caimo2017bayesian]. The typical criteria for interpreting the posteriors is that an odds ratio greater than one means a positive effect of a parameter, while odds ratios less than one represents negative effect. In addition, Bayesian approaches are also useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies. @Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on the average, 80% of the ties when a third of data missing.

@@

<!-- archaeological setting -->
We analyzed burial data collected from the published excavation reports, and the original fieldwork notes for the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD) [@Chen2007]. Excavations revealed abundant trade goods, including imported ceramics and stonewares, metal artifacts, glass beads, and stone beads, indicative of frequent trade activities [@Chen2007]. A total of 90 burials were unearthed that are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)). Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, suggesting a well organized spatial arrangement. Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences. For example, @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society. However, @Hsieh2012 suggested a relatively egalitarian structure based on comparative analysis of the frequencies of all burial goods. She found that the burials with high value goods were usually associated with elders, which might indicate achieved, rather than inherited, status. One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organization with post-European social organization. Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period to the post-European period. The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well studied.

@@

```{r sample-size}
# number of each phase
library(tidyverse)
pre_E <- burial_three_period_age_number %>% filter(Phase == "pre") %>% pull()
post_E <- burial_three_period_age_number %>% filter(Phase == "post") %>% pull()
chi <- burial_three_period_age_number %>% filter(Phase == "chi") %>% pull()
disturb <- burial_three_period_age_number %>% filter(Phase == "disturbed") %>% pull()
diff_pre_post <- post_E - pre_E
```

<!-- chronology -->
To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`). Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [@LWandBM2020ornament; @Chen2007]. We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size. There are `r disturb` burials heavily disturbed by modern construction that are also excluded because we cannot determine their chronology. Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf. @Coward2013, pp. 252], we built networks where burials (nodes in the network) that are linked when they have the same prestige goods in common. The prestige goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, and fish-shaped ornaments. Previous studies indicate that these items are considered as high-value across different archaeological contexts [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011]. Historical accounts written by Europeans and the Han Chinese also support these items as prestige goods [@Borao2009; @LiandWu2006].

@@

<!-- our hypothesis --> 
Our pericolonial impact hypothesis is that rare, high quality burial goods accumulated in only a few individuals' burials after the European presence because of increased social inequality at Kiwulan due to differential access to these goods. A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power, recognized as network strategies to achieve power [@Drennan2010; @Feinman2000]. If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity) and less transitivity (cohesion), along with greater differences in wealth accumulation according to burial goods. To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan. By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.

@@

```{r burial-value}
ave_value <- round(mean_burial_value,1) 
high_value <- round(high_burial_value,1) 
```

<!-- burial data and the definition of tie formation  --> 
We used burial goods in common to construct ties for representing a flow of good between burials. Since trade beads are commonly found across burials with substantial differences in quantities, we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials). For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting based on historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship. The networks constructed based on this principles show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:figure-network-diagrams)).

### 2021-01-26 analysis/paper/supplementary-material.Rmd
edit the paper rmd from intro to method

<!-- intro to networks: move to SI -->
A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994]. The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019]. For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp. 4]. Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts. Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013]. Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg. @Coward2013, pp. 252]. This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.

@@

<!-- current approaches to network analysis: move to SI -->
Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective ways to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020]. Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011]. For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby. Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics. By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007]. Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007]. ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.

>>>>

<!-- intro to networks: move to SI -->
A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994]. The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019]. For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp. 4]. Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts. Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013]. This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.

@@

<!-- current approaches to network analysis: move to SI -->
Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011]. For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby. Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics. By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007]. Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007]. ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
